---
title: "Analysis for primary lung function genes(T-paired) for github"
output: html_notebook
---



#load library
```{r}
library(dplyr)
library(tidyr)
library(writexl)
library(readxl)

library(purrr)

# set working directory
setwd(".")

rm(list=ls()) 
```

#Load data
```{r}
load("primary_genes/Exp_data_primary_CpG.RData")
load("Data/w0true.Rdata")
#W0.true is a matrix that represents all possible randomizations of paired data of 17 individuals

#let's call the final_primary_CpG to be final
final <- final_primary_CpG

```

#For primary gene list
```{r}
load("primary_genes/primary gene and CPG updated.RData") #this data default name is df_primary
df<-df_primary #call df_primary to be df instead

df1 <- df %>% select(1) 

row <-nrow(df) #number of CpG sites in primary gene list
nrow(w0.true)

```

#Section 1: calculate unadjusted fisher-exact p-values using paired |t|-statistics
```{r}
pair_idx <- split(seq_len(nrow(final)), final$id2)
n_pair <- length(pair_idx)
n_rand <- nrow(w0.true)

# methylation matrix: 34 observations x row CpG sites
Y <- as.matrix(final[, (4 + 1):(4 + row), drop = FALSE])

# numerator: mean paired difference under each randomization (all CpGs at once)
d_bar <- w0.true %*% Y / n_pair

# denominator: need pair-level squared differences for s_d
sum_D2 <- matrix(0, nrow = n_rand, ncol = row)
for (idx in pair_idx) {
  #for each pair, calculate the squared difference between the two subjects
  d_p <- w0.true[, idx, drop = FALSE] %*% Y[idx, , drop = FALSE]
  #sum the squared differences for all pairs
  sum_D2 <- sum_D2 + d_p^2
}

s_d <- sqrt((sum_D2 - n_pair * d_bar^2) / (n_pair - 1)) # sum_D2 - n_pair * d_bar^2 equals (\sum_i (d_i-\bar d)^2) in the formula in the paper
T_rand <- d_bar / (s_d / sqrt(n_pair))
T_rand_abs <- abs(T_rand)

# observed paired |t| for all CpGs
d_bar_obs <- as.numeric(crossprod(final$exp, Y) / n_pair)

# calculate the observed paired t for each pair 
# d_obs is 17*232 (17 pairs of differences with 232 CpGs on the columns)
d_obs <- t(vapply(
  pair_idx,
  function(idx) as.numeric(final$exp[idx] %*% Y[idx, , drop = FALSE]),
  numeric(row)
))
s_d_obs <- apply(d_obs, 2, sd) # calculate the standard deviation of the differences for each CpG; 2 means column; R default uses n-1 in the denominator for sample standard deviation for sd calculation 
T_obs_abs <- abs(d_bar_obs / (s_d_obs / sqrt(n_pair)))

ob_stat <- matrix(T_obs_abs, ncol = 1) # store the observed paired |t| for all CpGs
unadjust_p <- matrix(
  vapply(seq_len(row), function(h) mean(T_rand_abs[, h] >= T_obs_abs[h]), numeric(1)),
  ncol = 1
)

# store randomization statistics in list form (one vector per CpG)
test_stat <- lapply(seq_len(row), function(h) T_rand_abs[, h])

```

# Section 2: Adjust p-values by fully randomization based approach
# use paired |t| statistics from Section 1 (first 1000 randomizations)
```{r}
n <- 1000

test_stat1000 <- lapply(seq_len(row), function(h) T_rand_abs[1:n, h])
```


```{r}
#add all possible test statistics from each CpG together in the same data frame
all_stat1<- cbind(unlist(test_stat1000))

all_stat <- na.omit(all_stat1) #check for na values
  
k <- nrow(all_stat) #all_stat1 has k rows

```



```{r}
#create a matrix for all p values of all possible observed test statistics
p_matrix <- matrix(nrow=k, ncol=1)

#pretend every row in all_stat is observed stat, and calculate p value for each of them
for(i in 1:k) {
   #all_stat is all possible test statistics 
   b<- all_stat[i,1] #observed statistics
  p_value <- 1*sum(all_stat>=b)/k
  p_matrix[i,1] <- p_value
  print(i)
  }


#split the p-matrix by column cpg values, every 1000 test statistics corresponds to a CpG site
split_p <- split(p_matrix, ceiling(seq_along(p_matrix)/n)) #A one-liner splitting p_matrix into chunks of size n(number of rows in W0.rep)=1000:

z <- length(split_p)
```


```{r}
#store all minimum p values 
min <- matrix(nrow=n, ncol=1)

#store all corresponding elements in a matrix, and we will have 232=z matrixes in the end
mat1<- matrix(nrow=z , ncol=1)
mat <- list()

for (i in 1:n){ # i = randomization 1, 2, ..., 1000
  for (x in 1:z){ # x = CpG 1, 2, ..., 232
  mat1[x,1] <- split_p[[x]][i] # for ith randomization, we take all 232 p-values from each of 232 CpG sites
   }
  mat[[i]] <- mat1 #all 232 p-values from 232 CpGs at randomization i 
  min[i,1]<-min(mat[[i]]) #smallest p among those 232 for ith randomization 
}


#min(mat[[1]])

#split_p[[1]][3] #access the first element of the first list

```

```{r}
min <- as.data.frame(min)
```

# graph 
```{r}
library(esquisse)
#esquisser(min)
```

# graph distribution of minimum p-values
```{r}
ggplot(min) +
  aes(x = V1) +
  geom_histogram(bins = 93L, fill = "#5977AD") +
  labs(
    x = "Minimum p-values",
    title = "Distribution of \"minimum p-values\""
  ) +
  theme_linedraw()
```



# step 3 (calculate fully randomization based adjusted p-values): calculate the adjusted p values for the observed test statistics. All observed test statistics for each hypothesis are stored in ob_stat. Find the proportion of values in matrix min (minimum p value distribution) that are less than or equal to each value in the observed test statistics for each hypothesis, then that proportion is the adjusted p value
```{r}
# adjust p values for 232 hypothesis

adjust_p <- matrix(nrow=z, ncol=1)

for (i in 1:z){
  adjust_p[i, 1] <- 1*sum(min<=unadjust_p[i,1])/n #n is the number of values in the distribution of matrix min; z is the number of CpG sites in the primary gene list
}

adjust_p <- as.data.frame(adjust_p)

min(adjust_p)

min(unadjust_p)
```


# Section 3: FDR adjusted p-values for primary genes
```{r}
fdr_adjust <- as.data.frame (p.adjust(unadjust_p, method = "fdr", n = length(unadjust_p)))
fdr_adjust$fdr_adjusted <- fdr_adjust$`p.adjust(unadjust_p, method = "fdr", n = length(unadjust_p))`
fdr <- fdr_adjust %>% select(fdr_adjusted)
```



# Section 4: Caculate mean differences for primary genes' CpG sites
```{r}
mean_diff <- matrix(nrow=row, ncol=1)

#row represents the number of CpG sites in primary gene list 
for(h in 1:row){
  mean_diff[h,1] <-mean(final[,h+4][final$exp==1])-mean(final[,h+4][final$exp==-1])
  print(h)
}
```


# full join lung_genes with adjusted p value, and unadjusted p values
```{r}
lung_genes_primary_results <-
  adjust_p %>% cbind(df) %>% rename(ran_based_adjust_p=V1,CpG =Illumina, gene_name = X.19) %>%cbind(unadjust_p, fdr,mean_diff)  %>% select(2,3,5,1,6,7) %>% arrange(unadjust_p)

#optional: save adjusted p-value files
write_xlsx(lung_genes_primary_results,"primary_genes/lung_genes_primary_results.xlsx")
```


# Section 5: Visualizations
# vocalno plot of primary genes' unadjusted p-values
```{r}
library(ggplot2)
lung_genes_primary_results <- read_excel("primary_genes/lung_genes_primary_results.xlsx")

minimum <- 2/(2^17)

vn=expression(paste(-log[10],"(unadjusted p-values)"))
ti=expression(paste("Unadjusted Fisher-exact p-values of CpG sites of primary genes"))

label1=expression(paste(-log[10],(2/2^17))) ##

# create a graphs folder inside the primary_genes/code folder if it doesn't exist
if (!dir.exists("primary_genes/code/graphs")) {
  dir.create("primary_genes/code/graphs", showWarnings = FALSE)
}

pdf("primary_genes/code/graphs/prim_vol_unadjust.pdf", width = 7, height = 5, pointsize = 12, family = "Times")

volcano <- ggplot(data=lung_genes_primary_results, aes(x=mean_diff, y=-log10(unadjust_p))) + geom_point() + xlab("Average causal effect (ACE)") + ylab(vn) + xlim(-0.06, 0.06) + ylim(-0.3,5)+
geom_hline(yintercept=-log10(minimum), col="red")+annotate("text", x=0.053, y=5, label=label1, parse=TRUE)+   geom_point(data = lung_genes_primary_results[which.min(lung_genes_primary_results$unadjust_p), ],color="red", fill="red",shape=11, size=2) + theme_classic() #ggtitle(ti)

print(volcano)  # Ensure ggplot prints correctly before closing PDF
dev.off()

```

# volcano plot of primary genes' ran-based adjusted p-values
```{r}
min2 = 0.05
vn=expression(paste(-log[10],"(fully randomization-based adjusted p-values)"))
ti=expression(paste("Fully randomization-based adjusted p-values of CpG sites of primary genes"))

label1=expression(paste(-log[10],(0.05))) ##

pdf("primary_genes/code/graphs/prim_vol_rand_adj.pdf", width = 7, height = 5, pointsize = 12, family = "Times")
volcano2 <- ggplot(data=lung_genes_primary_results, aes(x=mean_diff, y=-log10(ran_based_adjust_p))) + geom_point() + xlab("Average causal effect (ACE)") + ylab(vn) + xlim(-0.06, 0.06) + ylim(-0.3,1.5)+
geom_hline(yintercept=-log10(min2), col="red")+annotate("text", x=0.055, y=1.4, label=label1)+   geom_point(data = lung_genes_primary_results[which.min(lung_genes_primary_results$ran_based_adjust_p), ],color="red", fill="red",shape=11, size=2) +theme_classic() # ggtitle(ti)


volcano2

dev.off()
```


# volcano plot of primary genes' FDR adjusted p-values
```{r}
vn=expression(paste(-log[10],"(FDR adjusted p-values)"))
ti=expression(paste("FDR adjusted p-values of CpG sites of primary genes"))

label1=expression(paste(-log[10],(0.05))) ##

pdf("primary_genes/code/graphs/prim_vol_fdr_adj.pdf", width = 7, height = 5, pointsize = 12, family = "Times")
volcano3 <- ggplot(data=lung_genes_primary_results, aes(x=mean_diff, y=-log10(fdr_adjusted))) + geom_point() + xlab("Average causal effect (ACE)") + ylab(vn) + xlim(-0.06, 0.06) + ylim(-0.3,1.5)+
geom_hline(yintercept=-log10(min2), col="red")+annotate("text", x=0.055, y=1.4, label=label1)+   geom_point(data = lung_genes_primary_results[which.min(lung_genes_primary_results$fdr_adjusted), ],color="red", fill="red",shape=11, size=2)+theme_classic() #ggtitle(ti)


volcano3
dev.off()
```

# histogram of all CpG sites of TLR2 gene in the primary gene group
```{r}
library(tidyverse)
primary_CpGs <- read_excel("primary_genes/primary gene and CPG.xlsx")

df <-
  primary_CpGs %>% filter(str_detect(X.19, "TLR2")) %>% select(1, 2) 

df 

row <- nrow(df)

pair_idx <- split(seq_len(nrow(final)), final$id2)
n_pair <- length(pair_idx)

paired_t_abs_for_cpg <- function(y) {
  D <- vapply(
    pair_idx,
    function(idx) as.numeric(w0.true[, idx, drop = FALSE] %*% y[idx]),
    numeric(nrow(w0.true))
  )
  d_bar <- rowMeans(D)
  s_d <- sqrt((rowSums(D^2) - n_pair * d_bar^2) / (n_pair - 1))
  T_rand_abs <- abs(d_bar / (s_d / sqrt(n_pair)))
  d_obs <- vapply(pair_idx, function(idx) sum(final$exp[idx] * y[idx]), numeric(1))
  T_obs_abs <- abs(mean(d_obs) / (sd(d_obs) / sqrt(n_pair)))
  list(null = T_rand_abs, obs = T_obs_abs)
}
```

# graph histogram for 12 cpg sites of gene TLR2 
```{r}
pdf("primary_genes/code/graphs/hist_TLR2.pdf", width = 9, height = 7, pointsize = 12, family = "Times")  

colnames(final) <- trimws(colnames(final))  # Remove leading/trailing spaces

par(mfrow = c(3, 4))
for(h in 1:row){
    #if else function, this column doesn't exist 
    #mn <- which(colnames(final)==df[h,1])
    j <- trimws(as.character(df[h,1]))  # Remove spaces
    if (j %in% names(final)) {
    #w0.true[1,]
    mn <- which(colnames(final)==j)
    stats <- paired_t_abs_for_cpg(as.numeric(final[, mn]))
    Test_Statistics <- stats$null
    hist(Test_Statistics, breaks = 100, main = paste("", df[h, 1]), xlab = "|T_paired|")
    abline(v = stats$obs, col = "red")
    }
    else {
      a<-NA
    }
}

dev.off()

```


# histogram of CpG sites with wider fisher intervals compared to their corresponding neymanian confidence intervals
# such CpG sites are \emph{cg05106269} of \emph{ICAM1}, \emph{cg02680732} of \emph{MUC5B}, \emph{cg16547110} of \emph{TLR2}, and \emph{cg10950028} of \emph{GSTM1} in the primary gene group
```{r}
# load library 
library(tidyverse)
# create graph folder inside the primary_genes/code folder if it doesn't exist
if (!dir.exists("primary_genes/code/graphs")) {
  dir.create("primary_genes/code/graphs", showWarnings = FALSE)
}
colnames(final) <- trimws(colnames(final))  # Remove leading/trailing spaces
primary_CpGs <- read_excel("primary_genes/primary gene and CPG.xlsx")

if (!exists("paired_t_abs_for_cpg")) {
  pair_idx <- split(seq_len(nrow(final)), final$id2)
  n_pair <- length(pair_idx)

  paired_t_abs_for_cpg <- function(y) {
    D <- vapply(
      pair_idx,
      function(idx) as.numeric(w0.true[, idx, drop = FALSE] %*% y[idx]),
      numeric(nrow(w0.true))
    )
    d_bar <- rowMeans(D)
    s_d <- sqrt((rowSums(D^2) - n_pair * d_bar^2) / (n_pair - 1))
    T_rand_abs <- abs(d_bar / (s_d / sqrt(n_pair)))
    d_obs <- vapply(pair_idx, function(idx) sum(final$exp[idx] * y[idx]), numeric(1))
    T_obs_abs <- abs(mean(d_obs) / (sd(d_obs) / sqrt(n_pair)))
    list(null = T_rand_abs, obs = T_obs_abs)
  }
}

```

## ICAM1 cg05106269
```{r}
# ICAM1
df_ICAM1 <-
  primary_CpGs %>% filter(str_detect(X.19, "ICAM1")) %>% select(1, 2) 

df_ICAM1 <- df_ICAM1 %>% filter(Illumina== "cg05106269")

print(df_ICAM1[1,1])

# create graph file inside the primary_genes/code/graphs folder
pdf(file.path("primary_genes/code/graphs", "null_rand_hist_ICAM1.pdf"), width = 9, height = 7, pointsize = 12, family = "Times")  

j <- trimws(as.character(df_ICAM1[1,1]))  # Remove spaces
if (j %in% names(final)) {
  mn <- which(colnames(final) == j)

  stats <- paired_t_abs_for_cpg(as.numeric(final[, mn]))
  Test_Statistics <- stats$null
  df_t <- n_pair - 1

  # density-scale histogram (comparable to theoretical |t| density)
  h <- hist(Test_Statistics, breaks = 100, plot = FALSE)
  bin_width <- diff(h$breaks)[1]
  h$density <- h$counts / (sum(h$counts) * bin_width)

  x_max <- round(max(Test_Statistics, stats$obs, qt(0.999, df = df_t)), 4)
  x_seq <- seq(0, x_max, length.out = 1000)
  y_asym <- 2 * dt(x_seq, df = df_t)
  y_max <- max(c(h$density, y_asym))
  
  # 1. Generate clean pretty ticks based on y_max
  y_ticks <- pretty(c(0, y_max), n = 6)
  
  # 2. Force the absolute maximum of those ticks to be our upper limit
  ylim_top <- max(y_ticks)

  old_mar <- par("mar")
  par(mar = c(8, old_mar[2], old_mar[3], old_mar[4]))
  plot(
    h,
    freq = FALSE,
    main = bquote(italic("ICAM1") ~ italic(.(as.character(df_ICAM1[1,1])))),
    xlab = "|T_paired|",
    xlim = c(0, x_max),
    ylab = "Density",
    ylim = c(0, ylim_top),
    col = "gray80",
    border = "white",
    xaxt = "n",
    yaxt = "n",
    xaxs = "i",
    yaxs = "r",
    cex.main = 1.5,
    cex.lab = 1.5
  )
  lines(x_seq, y_asym, col = "blue", lwd = 2)

  x_ticks <- pretty(c(0, x_max), n = 6)
  x_ticks <- sort(unique(c(x_ticks[x_ticks >= 0 & x_ticks < x_max], x_max)))
  axis(1, at = x_ticks, labels = FALSE)
  usr <- par("usr")
  text(
    x = x_ticks,
    y = usr[3] - 0.06 * diff(usr[3:4]),
    labels = formatC(x_ticks, digits = 8, format = "fg"),
    srt = 45,
    adj = 1,
    xpd = NA,
    cex = 1.2
  )

  axis(2, at = y_ticks, las = 1, labels = formatC(y_ticks, digits = 4, format = "fg"), xpd = NA, cex.axis = 1.5)
  box()
  par(mar = old_mar)

  abline(v = stats$obs, col = "red", lwd = 2)
  # 1. Force R to use flat, rectangular line caps ("butt" style)
  old_lend <- par(lend = 1)
  legend(
    "topright",
    legend = c(
      "Randomization |T_paired|",
      bquote("Asymptotic |t|, df =" ~ .(df_t)),
      "Observed |T_paired|"
    ),
    # Use lty = 1 for all items so R treats them as horizontal lines/bars
    lty = c(1, 1, 1), 
    col = c("gray80", "blue", "red"),
    # Set the gray bar to be very thick (lwd = 12), while keeping others at 2
    lwd = c(8, 2, 2), 
    bty = "n",
    # --- THESE PARAMETERS STRETCH THE MARKERS ---
    seg.len = 3.5,     # Stretches the horizontal width of the legend lines/bars
    x.intersp = 1.2,   # Adjusts text distance so it looks clean
    y.intersp = 1.3    # Adds vertical spacing between rows
  )
  # Restore the original line cap style so it doesn't affect other plots
  par(lend = old_lend)
} else {
  print("CpG site not found")
}


dev.off()
```



## MUC5B cg02680732
```{r}
# MUC5B
df_MUC5B <-
  primary_CpGs %>% filter(str_detect(X.19, "MUC5B")) %>% select(1, 2) 

df_MUC5B <- df_MUC5B %>% filter(Illumina== "cg02680732")

print(df_MUC5B[1,1])

# create graph file inside the primary_genes/code/graphs folder
pdf(file.path("primary_genes/code/graphs", "null_rand_hist_MUC5B.pdf"), width = 9, height = 7, pointsize = 12, family = "Times")  
j <- trimws(as.character(df_MUC5B[1,1]))  # Remove spaces
if (j %in% names(final)) {
  mn <- which(colnames(final) == j)

  stats <- paired_t_abs_for_cpg(as.numeric(final[, mn]))
  Test_Statistics <- stats$null
  df_t <- n_pair - 1

  h <- hist(Test_Statistics, breaks = 100, plot = FALSE)
  bin_width <- diff(h$breaks)[1]
  h$density <- h$counts / (sum(h$counts) * bin_width)

  x_max <- round(max(Test_Statistics, stats$obs, qt(0.999, df = df_t)), 4)
  x_seq <- seq(0, x_max, length.out = 1000)
  y_asym <- 2 * dt(x_seq, df = df_t)
  y_max <- max(c(h$density, y_asym))
  
  # 1. Generate clean pretty ticks based on y_max
  y_ticks <- pretty(c(0, y_max), n = 6)
  
  # 2. Force the absolute maximum of those ticks to be our upper limit
  ylim_top <- max(y_ticks)

  old_mar <- par("mar")
  par(mar = c(8, old_mar[2], old_mar[3], old_mar[4]))
  plot(
    h,
    freq = FALSE,
    main = bquote(italic("MUC5B") ~ italic(.(as.character(df_MUC5B[1,1])))),
    xlab = "|T_paired|",
    xlim = c(0, x_max),
    ylab = "Density",
    ylim = c(0, ylim_top),
    col = "gray80",
    border = "white",
    xaxt = "n",
    yaxt = "n",
    xaxs = "i",
    yaxs = "r",
    cex.main = 1.5,
    cex.lab = 1.5
  )
  lines(x_seq, y_asym, col = "blue", lwd = 2)

  x_ticks <- pretty(c(0, x_max), n = 6)
  x_ticks <- sort(unique(c(x_ticks[x_ticks >= 0 & x_ticks < x_max], x_max)))
  axis(1, at = x_ticks, labels = FALSE)
  usr <- par("usr")
  text(
    x = x_ticks,
    y = usr[3] - 0.06 * diff(usr[3:4]),
    labels = formatC(x_ticks, digits = 8, format = "fg"),
    srt = 45,
    adj = 1,
    xpd = NA,
    cex = 1.2
  )

  axis(2, at = y_ticks, las = 1, labels = formatC(y_ticks, digits = 4, format = "fg"), xpd = NA, cex.axis = 1.5)
  box()
  par(mar = old_mar)

  abline(v = stats$obs, col = "red", lwd = 2)
  old_lend <- par(lend = 1)
  legend(
    "topright",
    legend = c(
      "Randomization |T_paired|",
      bquote("Asymptotic |t|, df =" ~ .(df_t)),
      "Observed |T_paired|"
    ),
    lty = c(1, 1, 1),
    col = c("gray80", "blue", "red"),
    lwd = c(8, 2, 2),
    bty = "n",
    seg.len = 3.5,
    x.intersp = 1.2,
    y.intersp = 1.3
  )
  par(lend = old_lend)
} else {
  print("CpG site not found")
}

dev.off()
```

## TLR2 cg16547110
```{r}
# TLR2
df_TLR2 <-
  primary_CpGs %>% filter(str_detect(X.19, "TLR2")) %>% select(1, 2) 

df_TLR2 <- df_TLR2 %>% filter(Illumina== "cg16547110")

print(df_TLR2[1,1])

# create graph file inside the primary_genes/code/graphs folder
pdf(file.path("primary_genes/code/graphs", "null_rand_hist_TLR2.pdf"), width = 9, height = 7, pointsize = 12, family = "Times")  
j <- trimws(as.character(df_TLR2[1,1]))  # Remove spaces
if (j %in% names(final)) {
  mn <- which(colnames(final) == j)

  stats <- paired_t_abs_for_cpg(as.numeric(final[, mn]))
  Test_Statistics <- stats$null
  df_t <- n_pair - 1

  h <- hist(Test_Statistics, breaks = 100, plot = FALSE)
  bin_width <- diff(h$breaks)[1]
  h$density <- h$counts / (sum(h$counts) * bin_width)

  x_max <- round(max(Test_Statistics, stats$obs, qt(0.999, df = df_t)), 4)
  x_seq <- seq(0, x_max, length.out = 1000)
  y_asym <- 2 * dt(x_seq, df = df_t)
  y_max <- max(c(h$density, y_asym))
  
  # 1. Generate clean pretty ticks based on y_max
  y_ticks <- pretty(c(0, y_max), n = 6)
  
  # 2. Force the absolute maximum of those ticks to be our upper limit
  ylim_top <- max(y_ticks)
  # --- FIX END ---

  old_mar <- par("mar")
  par(mar = c(8, old_mar[2], old_mar[3], old_mar[4]))
  plot(
    h,
    freq = FALSE,
    main = bquote(italic("TLR2") ~ italic(.(as.character(df_TLR2[1,1])))),
    xlab = "|T_paired|",
    xlim = c(0, x_max),
    ylab = "Density",
    ylim = c(0, ylim_top),
    col = "gray80",
    border = "white",
    xaxt = "n",
    yaxt = "n",
    xaxs = "i",
    yaxs = "r",             # Change back to "r" so R leaves padding at the top for text
    cex.main = 1.5,
    cex.lab = 1.5
  )
  lines(x_seq, y_asym, col = "blue", lwd = 2)

  x_ticks <- pretty(c(0, x_max), n = 6)
  x_ticks <- sort(unique(c(x_ticks[x_ticks >= 0 & x_ticks < x_max], x_max)))
  axis(1, at = x_ticks, labels = FALSE)
  usr <- par("usr")
  text(
    x = x_ticks,
    y = usr[3] - 0.06 * diff(usr[3:4]),
    labels = formatC(x_ticks, digits = 8, format = "fg"),
    srt = 45,
    adj = 1,
    xpd = NA,
    cex = 1.2
  )

  axis(2, at = y_ticks, las = 1, labels = formatC(y_ticks, digits = 4, format = "fg"), xpd = NA, cex.axis = 1.5)
  box()
  par(mar = old_mar)

  abline(v = stats$obs, col = "red", lwd = 2)
  old_lend <- par(lend = 1)
  legend(
    "topright",
    legend = c(
      "Randomization |T_paired|",
      bquote("Asymptotic |t|, df =" ~ .(df_t)),
      "Observed |T_paired|"
    ),
    lty = c(1, 1, 1),
    col = c("gray80", "blue", "red"),
    lwd = c(8, 2, 2),
    bty = "n",
    seg.len = 3.5,
    x.intersp = 1.2,
    y.intersp = 1.3
  )
  par(lend = old_lend)
} else {
  print("CpG site not found")
}

dev.off()
```


## GSTM1 cg10950028
```{r}
# GSTM1
df_GSTM1 <-
  primary_CpGs %>% filter(str_detect(X.19, "GSTM1")) %>% select(1, 2) 

df_GSTM1 <- df_GSTM1 %>% filter(Illumina== "cg10950028")

print(df_GSTM1[1,1])

# create graph file inside the primary_genes/code/graphs folder
pdf(file.path("primary_genes/code/graphs", "null_rand_hist_GSTM1.pdf"), width = 9, height = 7, pointsize = 12, family = "Times")  
j <- trimws(as.character(df_GSTM1[1,1]))  # Remove spaces
if (j %in% names(final)) {
  mn <- which(colnames(final) == j)

  stats <- paired_t_abs_for_cpg(as.numeric(final[, mn]))
  Test_Statistics <- stats$null
  df_t <- n_pair - 1

  h <- hist(Test_Statistics, breaks = 100, plot = FALSE)
  bin_width <- diff(h$breaks)[1]
  h$density <- h$counts / (sum(h$counts) * bin_width)

  x_max <- round(max(Test_Statistics, stats$obs, qt(0.999, df = df_t)), 4)
  x_seq <- seq(0, x_max, length.out = 1000)
  y_asym <- 2 * dt(x_seq, df = df_t)
  y_max <- max(c(h$density, y_asym))
  
  # 1. Generate clean pretty ticks based on y_max
  y_ticks <- pretty(c(0, y_max), n = 6)
  
  # 2. Force the absolute maximum of those ticks to be our upper limit
  ylim_top <- max(y_ticks)

  old_mar <- par("mar")
  par(mar = c(8, old_mar[2], old_mar[3], old_mar[4]))
  plot(
    h,
    freq = FALSE,
    main = bquote(italic("GSTM1") ~ italic(.(as.character(df_GSTM1[1,1])))),
    xlab = "|T_paired|",
    xlim = c(0, x_max),
    ylab = "Density",
    ylim = c(0, ylim_top),
    col = "gray80",
    border = "white",
    xaxt = "n",
    yaxt = "n",
    xaxs = "i",
    yaxs = "r",
    cex.main = 1.5,
    cex.lab = 1.5
  )
  lines(x_seq, y_asym, col = "blue", lwd = 2)

  x_ticks <- pretty(c(0, x_max), n = 6)
  x_ticks <- sort(unique(c(x_ticks[x_ticks >= 0 & x_ticks < x_max], x_max)))
  axis(1, at = x_ticks, labels = FALSE)
  usr <- par("usr")
  text(
    x = x_ticks,
    y = usr[3] - 0.06 * diff(usr[3:4]),
    labels = formatC(x_ticks, digits = 8, format = "fg"),
    srt = 45,
    adj = 1,
    xpd = NA,
    cex = 1.2
  )

  axis(2, at = y_ticks, las = 1, labels = formatC(y_ticks, digits = 4, format = "fg"), xpd = NA, cex.axis = 1.5)
  box()
  par(mar = old_mar)

  abline(v = stats$obs, col = "red", lwd = 2)
  old_lend <- par(lend = 1)
  legend(
    "topright",
    legend = c(
      "Randomization |T_paired|",
      bquote("Asymptotic |t|, df =" ~ .(df_t)),
      "Observed |T_paired|"
    ),
    lty = c(1, 1, 1),
    col = c("gray80", "blue", "red"),
    lwd = c(8, 2, 2),
    bty = "n",
    seg.len = 3.5,
    x.intersp = 1.2,
    y.intersp = 1.3
  )
  par(lend = old_lend)
} else {
  print("CpG site not found")
}

dev.off()
```


# correlation matrix
```{r}
correlation_matrix <- final %>% filter(exp == -1) %>% select(c(5:236)) # keep only rows with clean air condition 
corr_prim <- cor(correlation_matrix) #correlation for all 688 variables

#example: correlation plot for first 30 CpG sites
pdf("primary_genes/code/graphs/corr_prim.pdf", width = 9, height = 7, pointsize = 12, family = "Times")  
correlation_matrix2 <- correlation_matrix %>% select(c(1:30))
library(corrplot)
source("http://www.sthda.com/upload/rquery_cormat.r")
rquery.cormat(correlation_matrix2, type="upper")
dev.off()
```


```{r}
library(rstatix)
#primary genes 
long_corr_prim <-  corr_prim %>% cor_gather() # long format of correlation matrix

group="Primary"

long_corr_prim <- long_corr_prim %>%cbind(group) #label the CpG with the group they belong to

#secondary genes 
load("secondary_genes/Exp_data_secondary_CpG.RData")
correlation_matrix_sec <- final_secondary_CpG %>% filter(exp == -1) %>% select(c(5:287))
corr_sec <- cor(correlation_matrix_sec) #correlation for all secondary CpG sites

long_corr_sec <-corr_sec %>% cor_gather() # long format of correlation matrix

group="Secondary"
long_corr_sec <- long_corr_sec %>%cbind(group) #label the CpG with the group they belong to
```



```{r}
#a table of all correlation values by groups

cor_by_group <- rbind(long_corr_prim,long_corr_sec)

cor_by_group <- cor_by_group %>% mutate(Sign = ifelse(cor >=0, "Positive", "Negative"))%>%rename("correlation"="cor")
```


# plot correlation for both primary and secondary genes
```{r}
library(ggplot2)

library(hrbrthemes)
library(viridis)

pdf("primary_genes/code/graphs/corr_violin_plot.pdf", width = 9, height = 7, pointsize = 12, family = "Times")  
p <- ggplot(cor_by_group, aes(x=group,y=correlation,fill=Sign))+ geom_boxplot(width=0.1)+ geom_violin(position="dodge",alpha=0.5,width=0.8,trim=TRUE)+theme_ipsum()+scale_fill_brewer(palette="Set1") + theme_classic()+xlab("Gene group")+ylab("Correlation")

p
dev.off()
```

```{r}
long_corr_prim %>% filter(cor < 0) %>% summary(cor)
long_corr_prim %>% filter(cor >0) %>% summary(cor)

long_corr_sec %>% filter(cor < 0) %>% summary(cor)
long_corr_sec%>% filter(cor >0) %>% summary(cor)
```

