---
title: "Fisherian interval primary genes graph"
output: html_notebook
---

#load library
```{r}
library(dplyr)
library(tidyr)

#clear the environment
rm(list=ls()) 

# set working directory to project root for reproducibility
setwd(".")
graph_dir <- "primary_genes/code/graphs"
if (!dir.exists(graph_dir)) {
  dir.create(graph_dir, recursive = TRUE, showWarnings = FALSE)
}

load("primary_genes/Exp_data_primary_CpG.RData")

load("Data/w0true.Rdata")
```

# Example 1: cg03610073 of TLR2. Effect size= -0.018702
```{r}
mn <- which(colnames(final_primary_CpG)=="cg03610073")   #provide the column number for the CpG site 
#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg03610073)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#constuct formulas to calculate p value for a list of a values and find the ones with p value larger than 0.025(due to two sided test(absolute value)) to construct fischer confidence interval

#for lower boundry. This CpG site has a negative effect size -->calculating p value should use <=
#should have no absolute value when calculating p value
```{r}
a=200
vectora_low = as.matrix(seq(-0.043, 0, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
print(i)
#when negative effect size, clean air (control) has larger values than the treatment O3
}
```


```{r}
library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_CA[w == -1], Yi_O3[w == 1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == -1], Y_obs[W_obs == 1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```


#for upper boundry. This CpG site has a negative effect size -->calculating p value should use >=
```{r}
a=200
vectora = as.matrix(seq(-0.035, 0, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_CA[w == -1], Yi_O3[w == 1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == -1], Y_obs[W_obs == 1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
# graphing of lower boundry (df1)
```{r}
pdf(file.path(graph_dir, "cg03610073_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

#lower boundry 
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df1 <- data.frame(vectora_low, list_p_low = list_p_low[,1])  # Convert to dataframe

ggplot(df1, aes(x = vectora_low, y = list_p_low)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg03610073",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
  
```
#graphing of upper boundry (df2)
```{r}
pdf(file.path(graph_dir, "cg03610073_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name
#upper boundry 
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df2 <- data.frame(vectora, list_p = list_p[,1])  # Convert to dataframe

ggplot(df2, aes(x = vectora, y = list_p)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg03610073",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
```
## fisherian interval is [-0.02953789,-0.0079]


# Section 2: cg18652319 of TLR2. effect size=0.002462
```{r}
mn <- which(colnames(final_primary_CpG)=="cg18652319")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg18652319)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```



#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=200
vectora_low = as.matrix(seq(0, 0.005, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1],paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1],paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```

#for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=200
vectora = as.matrix(seq(0, 0.006, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
# graphing of lower boundry (df3)
```{r}
pdf(file.path(graph_dir, "cg18652319_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

#lower boundry 
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df3 <- data.frame(vectora_low, list_p_low = list_p_low[,1])  # Convert to dataframe

ggplot(df3, aes(x = vectora_low, y = list_p_low)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg18652319",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
  
```

#graphing of upper boundry cg18652319 (df4)
```{r}
pdf(file.path(graph_dir, "cg18652319_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

#upper boundry 
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df4 <- data.frame(vectora, list_p = list_p[,1])  # Convert to dataframe

ggplot(df4, aes(x = vectora, y = list_p)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg18652319",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
```
# Confidence interval: [0.0004094737, 0.004467368]



#Section 3: cg03995486 of NOS2. effect sizes=0.008566
```{r}
mn <- which(colnames(final_primary_CpG)=="cg03995486")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg03995486)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=200
vectora_low = as.matrix(seq(-0.01, 0.025, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1],paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1],paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)


```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=200
vectora = as.matrix(seq(0, 0.028254847, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a
library(parallel)
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
#graphing of lower boundry (df5)
```{r}
#lower boundry 
library(ggplot2)
pdf(file.path(graph_dir, "cg03995486_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

# Assuming vectora_low and list_p_low are vectors
df5 <- data.frame(vectora_low, list_p_low = list_p_low[,1])  # Convert to dataframe

ggplot(df5, aes(x = vectora_low, y = list_p_low)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg03995486",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
```

#graphing of upper boundry (df6)
```{r}
#upper boundry 
pdf(file.path(graph_dir, "cg03995486_fish_int_up.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df6 <- data.frame(vectora, list_p = list_p[,1])  # Convert to dataframe

ggplot(df6, aes(x = vectora, y = list_p)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg03995486",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme
dev.off()
```
# 95% confidence interval is (0.001568421,0.01555263)


# Section 4: cg01133890 of MUC5B, effect size = -0.011
```{r}

mn <- which(colnames(final_primary_CpG)=="cg01133890")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg01133890)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a negative effect size -->calculating p value should use <=
#should have no absolute value when calculating p value
```{r}
a=200
vectora_low = as.matrix(seq(-0.03, 0, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_CA[w == -1], Yi_O3[w == 1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == -1], Y_obs[W_obs == 1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```

#for upper boundry. This CpG site has a negative effect size -->calculating p value should use >= 
```{r}
a=200
vectora = as.matrix(seq(-0.025, 0.008, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_CA[w == -1], Yi_O3[w == 1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == -1], Y_obs[W_obs == 1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
# graphing of lower boundry (df7)
```{r}
#lower boundry 
library(ggplot2)
pdf(file.path(graph_dir, "cg01133890_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

# Assuming vectora_low and list_p_low are vectors
df7 <- data.frame(vectora_low, list_p_low = list_p_low[,1])  # Convert to dataframe

ggplot(df7, aes(x = vectora_low, y = list_p_low)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg01133890",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
```

#graphing of upper boundry (df8)
```{r}
#upper boundry 
pdf(file.path(graph_dir, "cg01133890_fish_int_up.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df8 <- data.frame(vectora, list_p = list_p[,1])  # Convert to dataframe

ggplot(df8, aes(x = vectora, y = list_p)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg01133890",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme
dev.off()
```

# confidence interval is (-0.02094737, -0.002131579)

#Section 5: cg05106269 of ICAM1. effective size = 0.009854
```{r}
mn <- which(colnames(final_primary_CpG)=="cg05106269")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg05106269)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=200
vectora_low = as.matrix(seq(-0.015, 0.03, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1],paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1],paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=200
vectora = as.matrix(seq(-0.002631579, 0.03, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a
library(parallel)
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```

# visualization 
# graphing of lower boundry (df9)
```{r}
#lower boundry 
library(ggplot2)
pdf(file.path(graph_dir, "cg05106269_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

# Assuming vectora_low and list_p_low are vectors
df9 <- data.frame(vectora_low, list_p_low = list_p_low[,1])  # Convert to dataframe

ggplot(df9, aes(x = vectora_low, y = list_p_low)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg05106269",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
```

#graphing of upper boundry (df10)
```{r}
#upper boundry 
pdf(file.path(graph_dir, "cg05106269_fish_int_up.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df10 <- data.frame(vectora, list_p = list_p[,1])  # Convert to dataframe

ggplot(df10, aes(x = vectora, y = list_p)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg05106269",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme
dev.off()
```
# confidence interval is [0.0008894737,0.01916632]


#Section 6: cg00087425 of IL6. effect size = 0.0233. positive effect size 
```{r}
mn <- which(colnames(final_primary_CpG)=="cg00087425")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg00087425)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=200
vectora_low = as.matrix(seq(-0.025, 0.06, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1],paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1],paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=200
vectora = as.matrix(seq(-0.018, 0.06, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
# graphing of lower boundry (df11)
```{r}
#lower boundry 
library(ggplot2)
pdf(file.path(graph_dir, "cg00087425_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

# Assuming vectora_low and list_p_low are vectors
df11 <- data.frame(vectora_low, list_p_low = list_p_low[,1])  # Convert to dataframe

ggplot(df11, aes(x = vectora_low, y = list_p_low)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg00087425",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
```

#graphing of upper boundry (df12)
```{r}
#upper boundry 
pdf(file.path(graph_dir, "cg00087425_fish_int_up.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df12 <- data.frame(vectora, list_p = list_p[,1])  # Convert to dataframe

ggplot(df12, aes(x = vectora, y = list_p)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg00087425",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme
dev.off()
```

# confidence interval: [0.002226316, 0.04378947]


#Section 7: cg02680732 of MUC 5B. effect size = -0.003436
```{r}
#clear the environment
mn <- which(colnames(final_primary_CpG)=="cg02680732")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg02680732)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a negative effect size -->calculating p value should use <=
#should have no absolute value when calculating p value
```{r}
a=200
vectora_low = as.matrix(seq(-0.01052632, 0.003, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_CA[w == -1], Yi_O3[w == 1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == -1], Y_obs[W_obs == 1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```


#for upper boundry. This CpG site has a negative effect size -->calculating p value should use >=
```{r}
a=200
vectora = as.matrix(seq(-0.012, 0.003, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_CA[w == -1], Yi_O3[w == 1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == -1], Y_obs[W_obs == 1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
# graphing of lower boundry (df13)
```{r}
#lower boundry 
library(ggplot2)
pdf(file.path(graph_dir, "cg02680732_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

# Assuming vectora_low and list_p_low are vectors
df13 <- data.frame(vectora_low, list_p_low = list_p_low[,1])  # Convert to dataframe

ggplot(df13, aes(x = vectora_low, y = list_p_low)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect (a) for cg02680732",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
```

#graphing of upper boundry (df14)
```{r}
#upper boundry 
pdf(file.path(graph_dir, "cg02680732_fish_int_up.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df14 <- data.frame(vectora, list_p = list_p[,1])  # Convert to dataframe

ggplot(df14, aes(x = vectora, y = list_p)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg02680732",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme
dev.off()
```


# confidence interval: [-0.006736842, -0.0002347368]


#Section 8: cg16547110 of TLR2. effect size=0.005242
```{r}
#clear the environment

mn <- which(colnames(final_primary_CpG)=="cg16547110")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg16547110)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=200
vectora_low = as.matrix(seq(-0.005, 0.014, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1],paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1],paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```


# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=200
vectora = as.matrix(seq(-0.006, 0.015, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
# graphing of lower boundry (df15)
```{r}
#lower boundry 
library(ggplot2)
pdf(file.path(graph_dir, "cg16547110_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

# Assuming vectora_low and list_p_low are vectors
df15 <- data.frame(vectora_low, list_p_low = list_p_low[,1])  # Convert to dataframe

ggplot(df15, aes(x = vectora_low, y = list_p_low)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect (a) for cg16547110",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
```

#graphing of upper boundry (df16)
```{r}
#upper boundry 
pdf(file.path(graph_dir, "cg16547110_fish_int_up.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df16 <- data.frame(vectora, list_p = list_p[,1])  # Convert to dataframe

ggplot(df16, aes(x = vectora, y = list_p)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg16547110",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme
dev.off()
```

# confidence interval: [0.0001810526, 0.01021607]


#Section 9: cg10950028 of GSTM1. effect size = -0.0157
```{r}
mn <- which(colnames(final_primary_CpG)=="cg10950028")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg10950028)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a negative effect size -->calculating p value should use <=
#should have no absolute value when calculating p value
```{r}
a=200
vectora_low = as.matrix(seq(-0.053, 0.018421053, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_CA[w == -1], Yi_O3[w == 1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == -1], Y_obs[W_obs == 1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```


#for upper boundry. This CpG site has a negative effect size -->calculating p value should use >=
```{r}
a=200
vectora = as.matrix(seq(-0.06, 0.020526316, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_CA[w == -1], Yi_O3[w == 1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == -1], Y_obs[W_obs == 1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```

# confidence interval [-0.03226316, 0.0002789474]


# visualization 
# graphing of lower boundry (df17)
```{r}
#lower boundry 
library(ggplot2)
pdf(file.path(graph_dir, "cg10950028_fish_int_low.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")  # Specify the file name

# Assuming vectora_low and list_p_low are vectors
df17 <- data.frame(vectora_low, list_p_low = list_p_low[,1])  # Convert to dataframe

ggplot(df17, aes(x = vectora_low, y = list_p_low)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect (a) for cg10950028",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme

dev.off()
```

#graphing of upper boundry (df18)
```{r}
#upper boundry 
pdf(file.path(graph_dir, "cg10950028_fish_int_up.pdf"), width = 8, height = 6, pointsize = 12, family = "Times")
library(ggplot2)

# Assuming vectora_low and list_p_low are vectors
df18 <- data.frame(vectora, list_p = list_p[,1])  # Convert to dataframe

ggplot(df18, aes(x = vectora, y = list_p)) +
  geom_point(color = "blue", size = 2, alpha = 0.6) +  # Scatter plot
  geom_hline(yintercept = 0.025, linetype = "dashed", color = "red", size = 1) +  # Horizontal line
  labs(
    title = "P-values across different hypothetical constant treatment effect for cg10950028",
    x = "Hypothetical constant treatment effect (a)",
    y = "p-value"
  ) +
  theme_classic()   # Use a clean theme
dev.off()
```