---
title: "T paried fischer interval for selected primary genes (one example)"
output: html_notebook
---

#load library
```{r}
library(dplyr)
library(tidyr)

#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/primary_genes/Exp_data_primary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")
```

# Example 1: cg03610073 of TLR2. Effect size= -0.018702
```{r}
mn <- which(colnames(final_primary_CpG)=="cg03610073")   #provide the column number for the CpG site 
#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg03610073)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


# Calculate p value when the hypothesis is H_a=Yi(1)-Yi(0)=a, a is not equal to 0
```{r}
#trial "a" value
a=0.1

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but 0.009 more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+a
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-a

id <- final4$id2
W_obs <-final4$exp
Yi_wiO3<-final4$Yi_wiO3
Yi_wiCA<-final4$Yi_wiCA
Y_obs <- final4$Y_obs

stat <- matrix(ncol=1, nrow=n)

# t paired statistics
for(i in 1:n) {
  stat[i] <- t.test(Yi_wiO3[w0.true[i,]==1], Yi_wiCA[w0.true[i,]==-1],paired=TRUE)$statistic
}

stat

stat_obs <- t.test(Y_obs[W_obs==1],Y_obs[W_obs==-1], paired=TRUE)$statistic

stat_obs

#p-value calculation
p <- 1*sum(stat>=stat_obs)/n

p
```


#constuct formulas to calculate p value for a list of a values and find the ones with p value larger than 0.025(due to two sided test(absolute value)) to construct fischer confidence interval

#for lower boundry. This CpG site has a negative effect size -->calculating p value should use <=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(-0.02958, -0.02938, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4

#when negative effect size, clean air (control) has larger values than the treatment O3
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)

for(j in 1:n) {
  stat[j] <- t.test(Yi_CA[w0.true[j,]==-1],Yi_O3[w0.true[j,]==1], paired=TRUE)$statistic #we put clean air values first in t.test because it's expected to have larger values
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==-1], Y_obs[W_obs==1],paired=TRUE)$statistic

p_value <- 1*sum(stat<=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i, p_value))
}
```

#for upper boundry. This CpG site has a negative effect size -->calculating p value should use >=
```{r}
a=20
vectora = as.matrix(seq(-0.0089, -0.007, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_CA[w0.true[j,]==-1], Yi_O3[w0.true[j,]==1],  paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==-1], Y_obs[W_obs==1], paired=TRUE)$statistic

p_value <- 1*sum(stat>=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```
#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora, list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```
## fisherian interval is [-0.02953789,-0.0079]


# Section 2: cg18652319 of TLR2. effect size=0.002462
```{r}
#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/primary genes/Exp_data_primary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")

mn <- which(colnames(final_primary_CpG)=="cg18652319")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg18652319)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```



#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(0.00040, 0.00046, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- abs(t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic)
}

#observed statistics(no absolute value)
stat_obs <- abs(t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1], paired=TRUE)$statistic)

p_value <- 1*sum(stat>=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i,p_value))
}


```

#for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=20
vectora = as.matrix(seq(0.00444, 0.00448, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- abs(t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic)
}

#observed statistics(no absolute value)
stat_obs <- abs(t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic)

p_value <- 1*sum(stat<=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```


#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora, list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```
# Confidence interval: [0.0004094737, 0.004467368]

#Section 3: cg03995486 of NOS2. effect sizes=0.008566
```{r}
#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/primary genes/Exp_data_primary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")

mn <- which(colnames(final_primary_CpG)=="cg03995486")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg03995486)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(0.0013, 0.0016, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat>=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i,p_value))
}

```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=20
vectora = as.matrix(seq(0.0153, 0.0157, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat<=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```
# 95% confidence interval is (0.001568421,0.01555263)


# Section 4: cg01133890 of MUC5B, effect size = -0.011
```{r}
#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/primary genes/Exp_data_primary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")

mn <- which(colnames(final_primary_CpG)=="cg01133890")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg01133890)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a negative effect size -->calculating p value should use <=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(-0.022, -0.018, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_CA[w0.true[j,]==-1],Yi_O3[w0.true[j,]==1], paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==-1], Y_obs[W_obs==1],paired=TRUE)$statistic

p_value <- 1*sum(stat<=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i,p_value))
}

```

#for upper boundry. This CpG site has a negative effect size -->calculating p value should use >=
```{r}
a=20
vectora = as.matrix(seq(-0.0023, -0.0019, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_CA[w0.true[j,]==-1], Yi_O3[w0.true[j,]==1], paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==-1], Y_obs[W_obs==1],paired=TRUE)$statistic

p_value <- 1*sum(stat>=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```

# confidence interval is (-0.02094737, -0.002131579)

#Section 5: cg05106269 of ICAM1. effective size = 0.009854
```{r}
#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/primary genes/Exp_data_primary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")

mn <- which(colnames(final_primary_CpG)=="cg05106269")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg05106269)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(0.00080, 0.00097, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat>=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i,p_value))
}

```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=20
vectora = as.matrix(seq(0.01868, 0.01952, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat<=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```

# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```
# confidence interval is [0.0008894737,0.01916632]


#Section 6: cg00087425 of IL6. effect size = 0.0233. positive effect size 
```{r}
#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/primary genes/Exp_data_primary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")

mn <- which(colnames(final_primary_CpG)=="cg00087425")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg00087425)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(0.0021, 0.0025, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat>=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i,p_value))
}

```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=20
vectora = as.matrix(seq(0.043, 0.0455, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat<=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```

# confidence interval: [0.002226316, 0.04378947]


#Section 7: cg02680732 of MUC 5B. effect size = -0.003436
```{r}
#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/primary genes/Exp_data_primary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")

mn <- which(colnames(final_primary_CpG)=="cg02680732")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg02680732)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a negative effect size -->calculating p value should use <=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(-0.0068, -0.0065, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_CA[w0.true[j,]==-1],Yi_O3[w0.true[j,]==1], paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==-1], Y_obs[W_obs==1],paired=TRUE)$statistic

p_value <- 1*sum(stat<=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i,p_value))
}

```


#for upper boundry. This CpG site has a negative effect size -->calculating p value should use >=
```{r}
a=20
vectora = as.matrix(seq(-0.00026, -0.00018, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_CA[w0.true[j,]==-1], Yi_O3[w0.true[j,]==1], paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==-1], Y_obs[W_obs==1],paired=TRUE)$statistic

p_value <- 1*sum(stat>=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```

# confidence interval: [-0.006736842, -0.0002347368]


#Section 8: cg16547110 of TLR2. effect size=0.005242
```{r}
#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/primary genes/Exp_data_primary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")

mn <- which(colnames(final_primary_CpG)=="cg16547110")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg16547110)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(0.00016, 0.00021, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat>=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i,p_value))
}

```


# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=20
vectora = as.matrix(seq(0.01010526, 0.01031579, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat<=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```

# confidence interval: [0.0001810526, 0.01021607]


#Section 9: cg10950028 of GSTM1. effect size = -0.0157
```{r}
#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/primary genes/Exp_data_primary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")

mn <- which(colnames(final_primary_CpG)=="cg10950028")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_primary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg10950028)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a negative effect size -->calculating p value should use <=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(-0.033, -0.031, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_CA[w0.true[j,]==-1],Yi_O3[w0.true[j,]==1], paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==-1], Y_obs[W_obs==1],paired=TRUE)$statistic

p_value <- 1*sum(stat<=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i,p_value))
}

```

#for upper boundry. This CpG site has a negative effect size -->calculating p value should use >=
```{r}
a=20
vectora = as.matrix(seq(0.00015, 0.0005, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_CA[w0.true[j,]==-1], Yi_O3[w0.true[j,]==1], paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==-1], Y_obs[W_obs==1],paired=TRUE)$statistic

p_value <- 1*sum(stat>=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```

# confidence interval [-0.03226316, 0.0002789474]


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```