---
title: "Analysis for secondary lung function genes(T-paired) for github"
output: html_notebook
---


#load library
```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
library(writexl)
library(readxl)

library(purrr)

rm(list=ls()) 
```

#Load data
```{r}
load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/secondary genes/Exp_data_secondary_CpG.RData")
load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")
#W0.true is a matrix that represents all possible randomizations of paired data of 17 individuals

#let's call the final_secondary_CpG to be final
final <- final_secondary_CpG

```

#For secondary gene list
```{r}
load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/secondary genes/secondary gene and CPG.RData") #this data default name is df

row <-nrow(df)
nrow(w0.true)

```

#Section 1: calculate the observed unadjusted fisher-exact p-value for secondary genes
```{r}
test_stat <- list() #every CpG site has all possible 131072 test statistics
ob_stat <- matrix(nrow=row, ncol=1) #every CpG only has one observed statistics; so total of row=283 number of observed statistics for secondary gene list
unadjust_p <-matrix(nrow=row, ncol=1) #every CpG site only has one p value

for(h in 1:row){ #h represents the number of cpg column values
   test_stat[[h]] <-abs(w0.true%*%as.matrix(final[,h+4])/17)  #store all possible test statistics of each column inside a list of subjects
  ob_stat[h,1]<-abs(t(final$exp)%*%as.matrix(final[,h+4])/17) #store observed test statistics of each column
  unadjust_p[h,1]<-colSums((1*(sapply(seq_along(ob_stat[h,1]), function(x, y, i) y[i] <= x[,i], y = ob_stat[h,1] ,x = test_stat[[h]])))/2^17) #p-value, comparing one observed statistics (y) with all possible test statistics x for each CpG site
}

```

# Section 2: Adjust p-values by fully randomization based approach
# calculate all the test statistics for each CpG site using the first 1000 iterations of W0 true
```{r}
#we only consider the first 1000 randomizations. Select the first n rows of w0.true (randomization matrix)
n=1000
w0.true1 <- w0.true[c(1:n),]

test_stat1000 <- list() #every row has 1000 out of 131072 test statistics

for(h in 1:row){ 
  test_stat1000[[h]]<-abs(w0.true1%*%as.matrix(final[,h+4])/17)  #store all possible test statistics of each column inside a list of subjects
}
```


```{r}
#add all possible test statistics from each CpG together in the same data frame
all_stat1<- cbind(unlist(test_stat1000))

all_stat <- na.omit(all_stat1) #check for na values
  
k <- nrow(all_stat) #all_stat1 has k rows

```



```{r}
#create a matrix for all p values of all possible observed test statistics
p_matrix <- matrix(nrow=k, ncol=1)

#pretend every row in all_stat is observed stat, and calculate p value for each of them
for(i in 1:k) {
   #all_stat is all possible test statistics 
   b<- all_stat[i,1] #observed statistics
  p_value <- 1*sum(all_stat>=b)/k
  p_matrix[i,1] <- p_value
  print(i)
  }


#split the p-matrix by column cpg values, every 1000 test statistics corresponds to a CpG site
split_p <- split(p_matrix, ceiling(seq_along(p_matrix)/n)) #A one-liner splitting p_matrix into chunks of size n(number of rows in W0.rep)=1000:

z <- length(split_p)
```


```{r}
#store all minimum p values 
min <- matrix(nrow=n, ncol=1)

#store all corresponding elements in a matrix, and we will have 283=z matrixes in the end
mat1<- matrix(nrow=z , ncol=1)
mat <- list()

for (i in 1:n){ #n is number of elements in each list
   for (x in 1:z){ #z is number of lists
   mat1[x,1] <- split_p[[x]][i] #e.g. first elements of all lists will store in one list
   }
  mat[[i]] <- mat1 #end of step 2 where we make the list from the value from the same iteration of each list 
  min[i,1]<-min(mat[[i]]) #find the minimum of each list in mat
}


#min(mat[[1]])

#split_p[[1]][3] #access the first element of the first list

```

```{r}
min <- as.data.frame(min)
```

#graph
```{r}
library(esquisse)
esquisser(min)
```

#graph distribution of minimum p-values
```{r}
ggplot(min) +
  aes(x = V1) +
  geom_histogram(bins = 93L, fill = "#5977AD") +
  labs(
    x = "Minimum p-values",
    title = "Distribution of \"minimum p-values\""
  ) +
  theme_linedraw()
```



#step 3: (calculate fully randomization based adjusted p-values): calculate the adjusted p values for the observed test statistics. All observed test statistics for each hypothesis are stored in ob_stat. Find the proportion of values in matrix min (minimum p value distribution) that are less than or equal to each value in the observed test statistics for each hypothesis, then that proportion is the adjusted p value
```{r}
# adjust p values for 283 hypothesis

adjust_p <- matrix(nrow=x, ncol=1)

for (i in 1:x){
  adjust_p[i, 1] <- 1*sum(min<=unadjust_p[i,1])/n #n is the number of values in the distribution of matrix min
}

adjust_p <- as.data.frame(adjust_p)

min(adjust_p)

min(unadjust_p)
```


# Section 3: FDR adjusted p-values for secondary genes
```{r}
fdr_adjust <- as.data.frame (p.adjust(unadjust_p, method = "fdr", n = length(unadjust_p)))
fdr_adjust$fdr_adjusted <- fdr_adjust$`p.adjust(unadjust_p, method = "fdr", n = length(unadjust_p))`
fdr <- fdr_adjust %>% select(fdr_adjusted)
```


# Section 4: Caculate mean differences for secondary genes' CpG sites
```{r}
mean_diff <- matrix(nrow=row, ncol=1)

#row represents the number of CpG sites in primary gene list 
for(h in 1:row){
  mean_diff[h,1] <-mean(final[,h+4][final$exp==1])-mean(final[,h+4][final$exp==-1])
  print(h)
}
```

# full join lung_genes with adjusted p value, and unadjusted p values
```{r}
lung_genes_secondary_results <-
  adjust_p %>% cbind(df) %>% rename(ran_based_adjust_p=V1,CpG =Illumina, gene_name = X.19) %>%cbind(unadjust_p, fdr,mean_diff)  %>% select(2,3,5,1,6,7) %>% arrange(unadjust_p)

#optional: save adjusted p-value files
write_xlsx(lung_genes_secondary_results,"/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/secondary genes/lung_genes_secondary_results.xlsx")
```


# Section 5: Visualizations
# vocalno plot of primary genes' unadjusted p-values
```{r}
lung_genes_secondary_results <- read_excel("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/secondary_genes/lung_genes_secondary_results.xlsx")

minimum <- 2/(2^17)

vn=expression(paste(-log[10],"(unadjusted p-values)"))
ti=expression(paste("Unadjusted Fisher-exact p-values of CpG sites of secondary genes"))

label1=expression(paste(-log[10],(2/2^17))) ##

volcano <- ggplot(data=lung_genes_secondary_results, aes(x=mean_diff, y=-log10(unadjust_p))) + geom_point() + xlab("Average causal effect (ACE)") + ylab(vn) + xlim(-0.06, 0.06) + ylim(-0.3,5)+
geom_hline(yintercept=-log10(minimum), col="red")+annotate("text", x=0.053, y=5, label=label1)+   geom_point(data = lung_genes_secondary_results[which.min(lung_genes_secondary_results$unadjust_p), ],color="red", fill="red",shape=11, size=2) + ggtitle(ti)+theme_classic()


volcano

```
# vocalno plot of secondary genes' ran-based adjusted p-values
```{r}
min2 = 0.05
vn=expression(paste(-log[10],"(fully randomization-based adjusted p-values)"))
ti=expression(paste("Fully randomization-based adjusted p-values of CpG sites of secondary genes"))

label1=expression(paste(-log[10],(0.05))) ##

volcano2 <- ggplot(data=lung_genes_secondary_results, aes(x=mean_diff, y=-log10(ran_based_adjust_p))) + geom_point() + xlab("Average causal effect (ACE)") + ylab(vn) + xlim(-0.06, 0.06) + ylim(-0.3,1.5)+
geom_hline(yintercept=-log10(min2), col="red")+annotate("text", x=0.055, y=1.4, label=label1)+geom_point(data = lung_genes_secondary_results[which.min(lung_genes_secondary_results$ran_based_adjust_p), ],color="red", fill="red",shape=11, size=2) + ggtitle(ti)+theme_classic()


volcano2

```


# vocalno plot of primary genes' FDR adjusted p-values
```{r}
vn=expression(paste(-log[10],"(FDR adjusted p-values)"))
ti=expression(paste("FDR adjusted p-values of CpG sites of secondary genes"))

label1=expression(paste(-log[10],(0.05))) ##

minfdr <- lung_genes_secondary_results[1:5,]


volcano3 <- ggplot(data=lung_genes_secondary_results, aes(x=mean_diff, y=-log10(fdr_adjusted))) + geom_point() + xlab("Average causal effect (ACE)") + ylab(vn) + xlim(-0.06, 0.06) + ylim(-0.3,1.5)+
geom_hline(yintercept=-log10(min2), col="red")+annotate("text", x=0.055, y=1.4, label=label1)+   geom_point(data = minfdr,color="red", fill="red",shape=11, size=2) + ggtitle(ti)+theme_classic()


volcano3
```
