---
title: "Fisherian interval graph for important secondary genes"
output: html_notebook
---


#Section 1: cg12662193 of PLSCR1. effect sizes=0.00891
```{r}
#clear the environment
rm(list=ls()) 

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/secondary_genes/Exp_data_secondary_CpG.RData")

load("/Users/anqiwang/Documents/GitHub/Ozone-and-Lung-Functions/Data/w0true.Rdata")

mn <- which(colnames(final_secondary_CpG)=="cg12662193")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_secondary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg12662193)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(0.0042, 0.0046, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p_low <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat>=stat_obs)/n

list_p_low[i,1]<-p_value

print(c(i,p_value))
}

```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=10
vectora = as.matrix(seq(0, 0.019, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

list_p <- matrix(nrow=a, ncol=1)     # nrow= number of a values

for(i in 1:a) {
id <- x[[i]]$"id2"
W_obs <-x[[i]]$"exp"
Yi_CA<-x[[i]]$"Yi_wiCA"
Yi_O3<-x[[i]]$"Yi_wiO3"
Y_obs <- x[[i]]$"Y_obs"
stat <- matrix(ncol=1, nrow=n)


for(j in 1:n) {
  stat[j] <- t.test(Yi_O3[w0.true[j,]==1], Yi_CA[w0.true[j,]==-1],paired=TRUE)$statistic
}

#observed statistics(no absolute value)
stat_obs <- t.test(Y_obs[W_obs==1], Y_obs[W_obs==-1],paired=TRUE)$statistic

p_value <- 1*sum(stat<=stat_obs)/n

list_p[i,1]<-p_value

print(c(i, p_value))
}
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```

# confidence CI: [0.004473684, 0.01342105]



# section 2: cg19869469 of C6orf227. effect size= 0.017787267
```{r}
mn <- which(colnames(final_secondary_CpG)=="cg19869469")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_secondary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg19869469 )

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=15
vectora_low = as.matrix(seq(-0.01, 0.033, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1],paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1],paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```


# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=20
vectora = as.matrix(seq(0, 0.031315789, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a
library(parallel)
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```

# confidence interval: [0.009368421, 0.02615236]



# section 3: cg14795253 of PLSCR1. effect size = 0.006089086
```{r}
mn <- which(colnames(final_secondary_CpG)=="cg14795253")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_secondary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg14795253)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(-0.004, 0.013, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1],paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1],paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=20
vectora = as.matrix(seq(0.007, 0.012, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a
library(parallel)
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```


# confidence interval: [0.002891579, 0.009277008]


# section 4: cg08301503 of C6orf227. effect size = 0.020336783
```{r}
mn <- which(colnames(final_secondary_CpG)=="cg08301503")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_secondary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg08301503)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(-0.01, 0.047368421, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a

library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1],paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1],paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=20
vectora = as.matrix(seq(-0.01, 0.047368421, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a
library(parallel)
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```


# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```

# confidence interval: [0.007710526, 0.03286842]


# section 5: cg01392313 of C6orf227. effect size = 0.028433911
```{r}
mn <- which(colnames(final_secondary_CpG)=="cg01392313")   #provide the column number for the CpG site 

#all possible randomizations
n <- 2^17

#create a data frame that contains the CpG site of interest
final1<-
  final_secondary_CpG %>%
  select(c(1:4),mn) %>%
  rename(Y_obs = cg01392313)

#data wrangling
#ozone, 1=O3
final2 <-
  final1%>%
  filter(exp==1) %>%
  rename(Yi_wiO3 = Y_obs)%>%
  select(3,4,5)

#clean air, -1=CA
final3 <-
  final1 %>%
  filter(exp==-1)%>%
  rename(Yi_wiCA = Y_obs)%>%
  select(3,4,5)


#full join table final4 with final2 or final 3, respectively
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)
```


#for lower boundry. This CpG site has a positive effect size -->calculating p value should use >=
#should have no absolute value when calculating p value
```{r}
a=20
vectora_low = as.matrix(seq(-0.012, 0.06, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora_low[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora_low[i,1]
x[[i]]<-final4
}
```


```{r}
library(parallel)

list_p_low <- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1],paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1],paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat>=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p_low[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)

```

# for upper boundry. This CpG site has a positive effect size -->calculating p value should use <=
```{r}
a=20
vectora = as.matrix(seq(-0.015526316, 0.060526316, length.out = a)) #if changing the number of a values here length.out=# of a values

x<-list()

for(i in 1:a) { # of a value in the loop
  #substitute the NA value in Yi_1 with the correponding value in Yi_0 but "a" more
final4 <-
  full_join(final1, final2, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,6,5)

final4<-
  full_join(final4, final3, by = c("id2" = "id2", "exp" = "exp")) %>%
  select(1,2,3,4,5,7,6)

#substitute the NA value in Yi_1 with the correponding value in Yi_0 but a more
final4$Yi_wiO3[is.na(final4$Yi_wiO3)] <- final4$Yi_wiCA[final4$exp==-1]+vectora[i,1]
final4$Yi_wiCA[is.na(final4$Yi_wiCA)] <- final4$Yi_wiO3[final4$exp==1]-vectora[i,1]
x[[i]]<-final4
}
```


```{r}
#create a matrix with all p-values corresponding to each value a
library(parallel)
list_p<- matrix(nrow=a, ncol=1)  # nrow= number of a values

# Function for processing a single `i`
compute_p_value <- function(i) {
  id <- x[[i]]$"id2"
  W_obs <- x[[i]]$"exp"
  Yi_CA <- x[[i]]$"Yi_wiCA"
  Yi_O3 <- x[[i]]$"Yi_wiO3"
  Y_obs <- x[[i]]$"Y_obs"

  # Compute statistics for all j in one step
  stat <- apply(w0.true, 1, function(w) { #1 represents applying each row in W0.true
    t.test(Yi_O3[w == 1], Yi_CA[w == -1], paired=TRUE)$statistic
  })

  # Observed statistics (no absolute value)
  stat_obs <- t.test(Y_obs[W_obs == 1], Y_obs[W_obs == -1], paired=TRUE)$statistic

  # Compute p-value
  p_value <- 1*sum(stat<=stat_obs)/n
  
  return(p_value)
}

# Parallel execution
cl <- makeCluster(detectCores() - 1)  # Use all but one core
clusterExport(cl, c("x", "w0.true", "n"))  # Added "n" to the export list
library(pbapply)
list_p[, 1] <- pbsapply(1:a, compute_p_value, cl = cl) # each run compute_p_value(i) for different i values.
stopCluster(cl)
```

# confidence interval: [0.01205263, 0.04490304]

# visualization 
# graphing of lower boundry 
```{r}
#lower boundry 
plot(vectora_low, list_p_low)
abline(h=0.025)

#we can see that the lower boundry corresponds to a= -0.02953789
```

#graphing of upper boundry
```{r}
#upper boundry 
plot(vectora,list_p)
abline(h=0.025)

#we can see that the upper boundry corresponds to a= -0.0079
```
